---
title: "College Softball Batting Analysis: 2021 vs. 2022"
author: "Emma Caetano"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(tidyverse)

# Read 2021 and 2022 batting datasets and label the season
batting_2021 <- read_csv("https://data.scorenetwork.org/data/batting2021.csv") |>
  mutate(year = "2021")

batting_2022 <- read_csv("https://data.scorenetwork.org/data/batting2022.csv") |>
  mutate(year = "2022")

# Combine both seasons into a single dataset
batting <- bind_rows(batting_2021, batting_2022) |>
  filter(AB >= 50)
```

## Comparing On-Base Plus Slugging (OPS) Across Seasons

```{r}
#| label: ops-comparison-plot
#| fig-cap: "Distribution of On-Base Plus Slugging (OPS) for NCAA softball players with at least 50 at-bats in 2021 and 2022."

ggplot(batting, aes(x = OPS, fill = year)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("2021" = "#1f77b4", "2022" = "#ff7f0e")) +
  labs(
    title = "NCAA Softball Batting Performance (2021 vs. 2022)",
    subtitle = "Comparing On-Base Plus Slugging (OPS) distributions across both seasons",
    x = "On-Base Plus Slugging (OPS)",
    y = "Density",
    fill = "Season"
  ) +
  theme_minimal()
```

Across the 2021 and 2022 college softball seasons, overall offensive production remained remarkably consistent among qualified hitters (players with at least 50 at-bats). The average OPS was 0.787 in 2021 and 0.783 in 2022, with both seasons displaying nearly identical unimodal distributions centered around a median OPS of approximately 0.76. While general offensive rates such as batting average and on-base percentages held steady, power hitting saw a slight uptick in 2022, with players averaging 4.06 home runs compared to 3.61 in 2021.
